\documentclass[aspectratio=169,10pt]{beamer}

\usetheme{Madrid}
\usecolortheme{default}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{itemize items}[circle]

\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage[T1]{fontenc}
\usepackage{array}

% Table directories
\newcommand{\tfirm}{../tables/firm}
\newcommand{\tbndes}{../tables/agg_firm_bndes_sector}

% Colors
\definecolor{darkblue}{RGB}{0,51,102}
\definecolor{alertred}{RGB}{180,30,30}
\definecolor{pass}{RGB}{198,232,198}

\setbeamercolor{frametitle}{fg=white}
\setbeamercolor{title}{fg=white}
\setbeamercolor{structure}{fg=darkblue}

\newcommand{\fpass}[1]{\cellcolor{pass}\textbf{#1}}

\title{First-Stage Summary: Key Findings}
\subtitle{Scripts 51 \& 52 --- Abril 17, 2026}
\author{}
\date{}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

% ===========================================================================
% MAIN FINDINGS
% ===========================================================================

\begin{frame}[t]{Main Findings}
\begin{itemize}\setlength{\itemsep}{0.75em}

  \item \textbf{Emp-share weighting is relevant and improves on emp-weighting
    --- for the president instrument only.}
    At the firm level, the presidential alignment instrument ($P$) rises from
    $F=15$ to $F=38$ (cycle-specific) and from $F=5.5$ to $F=33$ (2002-fixed)
    when switching from emp-weighted to emp-share-weighted.  Mayor and governor
    instruments remain dominated by the unweighted spec ($F>45$, $F>100$).
    At the \emph{aggregated} level, emp-share-weighted produces numerical
    artifacts (near-singular cells) and cannot be used.

  \item \textbf{Splitting the firm sample reveals no concentration: both
    subsamples have a strong first stage.}
    For BNDES extensive (cycle-specific, unweighted, coalition, pooled),
    the governor instrument yields $F=64$ in top-quartile municipalities and
    $F=61$ in the bottom three quartiles.  The first-stage signal at the firm
    level is broad, not city-size-specific.

  \item \textbf{Splitting the \emph{aggregated} sample reveals sharp
    concentration in large cities.}
    With muni$\times$year FE, the governor instrument reaches $F=23.5$ in
    top-quartile municipalities but only $F=0.4$ in the rest
    (full-sample $F$ was $2.2$ --- diluted by small cities).
    For party alignment the joint M+G+P combo gives $F=14.6$ in large cities
    vs.\ $F=1.3$ elsewhere.  The sector-level first stage is viable only for
    top-quartile municipalities.

\end{itemize}
\end{frame}

% ===========================================================================
% F-GRID: WEIGHTING COMPARISON — FIRM LEVEL
% ===========================================================================

\begin{frame}[t]{Firm Level: Weighting Comparison --- BNDES Extensive, Coalition, Pooled-count}
\framesubtitle{All firms $\cdot$ Coalition alignment $\cdot$ Pooled-count exposure}
\begin{center}
\footnotesize
\setlength{\tabcolsep}{3.5pt}
\renewcommand{\arraystretch}{1.05}
\begin{tabular}{@{}llcccccc@{}}
\toprule
 & & \multicolumn{6}{c}{Instrument combo} \\
\cmidrule(l{2pt}r{2pt}){3-8}
Baseline & Weighting & M & G & P & M+G & M+P & M+G+P \\
\midrule
\multirow{3}{*}{Cycle-specific}
  & Unweighted      & \fpass{45.5} & \fpass{103.2} & 0.9  & \fpass{63.3} & \fpass{22.9} & \fpass{43.4} \\
  & Emp-weighted    & 0.2          & 0.9           & \fpass{15.0}  & 0.5          & 7.9          & 5.9          \\
  & Emp-share-wt'd  & 0.4          & 0.8           & \fpass{38.2}  & 0.5          & \fpass{11.0} & 7.6          \\
\addlinespace[5pt]
\multirow{3}{*}{2002-fixed}
  & Unweighted      & \fpass{13.7} & \fpass{23.7}  & \fpass{24.4} & \fpass{16.3} & \fpass{20.2} & \fpass{22.6} \\
  & Emp-weighted    & 0.2          & 2.5           & 5.5          & 1.4          & 2.9          & 4.1          \\
  & Emp-share-wt'd  & 1.2          & 2.7           & \fpass{33.2} & 1.2          & 9.7          & 7.2          \\
\bottomrule
\end{tabular}
\end{center}
{\scriptsize \colorbox{pass}{\textbf{Shaded}} = $F>10$.
Emp-share-wt'd improves strongly on emp-weighted for $P$ (president), but unweighted dominates for $M$ (mayor) and $G$ (governor).
Firm + muni$\times$year FE. SEs clustered by firm + muni.}
\end{frame}

% ===========================================================================
% F-GRID: SPLIT SAMPLE — FIRM LEVEL
% ===========================================================================

\begin{frame}[t]{Firm Level: Split Sample --- BNDES Extensive, Coalition, Cycle-specific, Unweighted}
\framesubtitle{Pooled-count exposure $\cdot$ Main instrument family}
\begin{center}
\footnotesize
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.05}
\begin{tabular}{@{}lcccccc@{}}
\toprule
 & \multicolumn{6}{c}{Instrument combo} \\
\cmidrule(l{2pt}r{2pt}){2-7}
Sample & M & G & P & M+G & M+P & M+G+P \\
\midrule
All firms  & \fpass{45.5} & \fpass{103.2} & 0.9 & \fpass{63.3} & \fpass{22.9} & \fpass{43.4} \\
Top Q4     & \fpass{21.6} & \fpass{63.9}  & 1.3 & \fpass{37.1} & \fpass{11.3} & \fpass{24.9} \\
Bottom 3Q  & \fpass{33.3} & \fpass{61.1}  & 0.1 & \fpass{40.6} & \fpass{16.8} & \fpass{28.8} \\
\bottomrule
\end{tabular}
\end{center}
{\scriptsize \colorbox{pass}{\textbf{Shaded}} = $F>10$.
Both subsamples show a strong first stage.  Top Q4: 37.9M obs; Bottom 3Q: 5.2M obs.
Firm + muni$\times$year FE. SEs clustered by firm + muni.}
\vspace{0.8em}

\textcolor{alertred}{\footnotesize\textbf{Implication:} the alignment instrument is a valid first stage in large and small cities alike at the firm level. City-size heterogeneity arises only at the aggregated (sector) level (next slide).}
\end{frame}

% ===========================================================================
% F-GRID: SPLIT SAMPLE — AGGREGATED
% ===========================================================================

\begin{frame}[t]{Aggregated Level: Split Sample --- BNDES Extensive, Unweighted, Pooled-count, Cycle-specific}
\framesubtitle{BNDES sector $\cdot$ Owner-count aggregation $\cdot$ Exposure control}
\begin{center}
\footnotesize
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.05}
\begin{tabular}{@{}lllcccccc@{}}
\toprule
 & & & \multicolumn{6}{c}{Instrument combo} \\
\cmidrule(l{3pt}r{3pt}){4-9}
Alignment & FE & Sample & M & G & P & M+G & M+P & M+G+P \\
\midrule
\multirow{4}{*}{Coalition}
 & \multirow{2}{*}{MxJ+JxT}
   & Top Q4    & 1.2 & 2.6           & 1.6 & 0.8           & 1.0 & 0.7           \\
 & & Bottom 3Q & 0.0 & 0.3           & 0.3 & 0.2           & 0.9 & 0.8           \\
 & \multirow{2}{*}{MxJ+MxT}
   & Top Q4    & 1.2 & \fpass{23.5}  & 1.0 & \fpass{16.1}  & 0.4 & \fpass{13.4}  \\
 & & Bottom 3Q & 0.1 & 0.4           & 1.5 & 0.2           & 0.7 & 0.5           \\
\addlinespace[4pt]
\multirow{4}{*}{Party}
 & \multirow{2}{*}{MxJ+JxT}
   & Top Q4    & 1.4 & 2.4           & 3.3 & 1.6           & 6.0 & \fpass{14.6}  \\
 & & Bottom 3Q & 2.9 & 1.4           & 0.0 & 0.8           & 0.7 & 1.3           \\
 & \multirow{2}{*}{MxJ+MxT}
   & Top Q4    & 0.3 & 0.0           & 2.3 & 0.2           & 1.2 & 1.1           \\
 & & Bottom 3Q & 0.0 & 0.5           & 2.5 & 0.1           & 1.2 & 0.9           \\
\bottomrule
\end{tabular}
\end{center}
{\scriptsize \colorbox{pass}{\textbf{Shaded}} = $F>10$.
Full-sample coalition MxJ+MxT: G=2.2, M+G=3.0, M+G+P=3.3 (all weak --- diluted by small cities).
Muni$\times$sector + year FE. SEs clustered by muni + sector.}
\end{frame}

% ===========================================================================
% SELECTED REGRESSION TABLES
% ===========================================================================

\begin{frame}[plain]
\begin{center}
{\Large\textbf{Selected Regression Tables}}
\end{center}
\end{frame}

% --- Firm: best unweighted spec ---------------------------------------------

\begin{frame}[t,shrink=5]{Firm: BNDES Extensive --- Coal $\cdot$ Cyc $\cdot$ Unweighted $\cdot$ All firms}
\begin{center}\footnotesize
\input{\tfirm/firm__main__levels__bndes_extensive__coalition__cycle_specific__unweighted__all_firms__pooled_count}
\end{center}
\end{frame}

% --- Firm: best emp-share-weighted spec -------------------------------------

\begin{frame}[t,shrink=5]{Firm: BNDES Extensive --- Coal $\cdot$ Cyc $\cdot$ Emp-share-weighted $\cdot$ All firms}
\begin{center}\footnotesize
\input{\tfirm/firm__main__levels__bndes_extensive__coalition__cycle_specific__emp_share_weighted__all_firms__pooled_count}
\end{center}
\end{frame}

% --- Firm split: top Q4 vs bottom 3Q ----------------------------------------

\begin{frame}[t,shrink=5]{Firm Split: BNDES Extensive --- Top-Quartile Municipalities}
\framesubtitle{Cycle-specific $\cdot$ Unweighted $\cdot$ Coalition $\cdot$ Pooled-count}
\begin{center}\footnotesize
\input{\tfirm/firm__main__levels__bndes_extensive__coalition__cycle_specific__unweighted__top_q4__pooled_count}
\end{center}
\end{frame}

\begin{frame}[t,shrink=5]{Firm Split: BNDES Extensive --- Bottom Three Quartiles}
\framesubtitle{Cycle-specific $\cdot$ Unweighted $\cdot$ Coalition $\cdot$ Pooled-count}
\begin{center}\footnotesize
\input{\tfirm/firm__main__levels__bndes_extensive__coalition__cycle_specific__unweighted__bottom_3q__pooled_count}
\end{center}
\end{frame}

% --- Aggregated split: coalition MxJ+MxT -----------------------------------

\begin{frame}[t,shrink=5]{Agg Split: BNDES Extensive, Coalition, MxJ+MxT --- Top-Quartile Municipalities}
\framesubtitle{BNDES sector $\cdot$ Unweighted $\cdot$ Cycle-specific $\cdot$ Pooled-count}
\begin{center}\footnotesize
\input{\tbndes/agg_firm__bndes_sector__bndes_extensive__coalition__cycle_specific__owner_count__unweighted__mxj_mxt__ctrl__pooled_count__top_q4}
\end{center}
\end{frame}

\begin{frame}[t,shrink=5]{Agg Split: BNDES Extensive, Coalition, MxJ+MxT --- Bottom Three Quartiles}
\framesubtitle{BNDES sector $\cdot$ Unweighted $\cdot$ Cycle-specific $\cdot$ Pooled-count}
\begin{center}\footnotesize
\input{\tbndes/agg_firm__bndes_sector__bndes_extensive__coalition__cycle_specific__owner_count__unweighted__mxj_mxt__ctrl__pooled_count__bottom_3q}
\end{center}
\end{frame}

\end{document}
