\documentclass{article}
\usepackage[top=2.3cm, bottom=2.3cm, left=2.1cm, right=2.1cm]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage[default,regular,black]{sourceserifpro}
\usepackage[T1]{fontenc}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}

\newcommand{\FA}{\mathrm{FA}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cFpre}{\mathcal{F}^{\mathrm{pre}}}

\theoremstyle{plain}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\theoremstyle{definition}
\newtheorem{condition}{Condition}
\newtheorem{example}{Example}

\begin{document}

\title{Understanding Conditions C3, C5, and C6\\
for Proposition~2 Aggregation Equivalence}
\author{}
\date{\today}
\maketitle

\section*{Recap: The Setup}

The firm-level regression is
\begin{equation}\label{eq:firm}
  \ind(\text{BNDES}_{fmt} > 0)
  = \sum_\ell \lambda_\ell\, \FA^\ell_{fmt}
  + \gamma_f + \alpha_{mt} + u_{fmt},
\end{equation}
where $f$ indexes firms, $m$ municipalities, $t$ years, and $\ell$ alignment
tiers.  The fixed effects are $\gamma_f$ (firm) and $\alpha_{mt}$
(municipality$\times$year).

Each firm belongs to a \textbf{cell} $c = (j, m, e)$, where $j$ is the sector,
$m$ is the municipality, and $e$ is the election cycle (support regime).  The
pre-election firm set in cell $c$ is $\cFpre_c$ with $N_c = |\cFpre_c|$ firms.

Averaging equation~\eqref{eq:firm} over firms in cell $c$ gives the
\textbf{cell-level equation}:
\begin{equation}\label{eq:cell}
  H^\mathrm{pre}_{jmt}
  = \sum_\ell \lambda_\ell\,
    \overline{\FA}^{\ell,\mathrm{pre}}_{jmt}
  + \bar\gamma_c + \alpha_{mt} + \epsilon_{jmt},
\end{equation}
where
$H^\mathrm{pre}_{jmt} = N_c^{-1}\sum_{f \in \cFpre_c} \ind(\text{BNDES}_{fmt} > 0)$,
$\overline{\FA}^{\ell,\mathrm{pre}}_{jmt} = N_c^{-1}\sum_{f \in \cFpre_c} \FA^\ell_{fmt}$,
$\bar\gamma_c = N_c^{-1}\sum_{f \in \cFpre_c} \gamma_f$, and
$\epsilon_{jmt} = N_c^{-1}\sum_{f \in \cFpre_c} u_{fmt}$.

Proposition~2 claims that running OLS on~\eqref{eq:firm} (unweighted) and
running $N_c$-weighted OLS on~\eqref{eq:cell} produce the \textbf{same}
$\hat\lambda$.  This requires conditions C1--C6.  Below we explain C3, C5, and
C6 in detail.

\bigskip
\hrule
\bigskip

%% ========================================================================
\section{Condition C3: Firm Immobility}
%% ========================================================================

\begin{condition}[C3 --- Firm Immobility]
Each firm $f$ belongs to exactly one cell $(j, m)$ across all years in the
sample.  Formally:
\[
  \forall\, f:\quad
  \bigl|\{(j, m) : \exists\, t \text{ with } (f, j, m, t) \in \text{sample}\}\bigr| = 1.
\]
\end{condition}

\subsection{Intuitive Explanation}

Imagine you sort kids into classrooms based on their neighborhood and favorite
subject.  Each kid has a ``personal talent score'' ($\gamma_f$) that stays with
them forever.

If every kid stays in the same classroom for the whole school, you can compute
each classroom's ``average talent'' and use it as a single classroom label.  No
information leaks between classrooms.

But suppose Alice moves from Classroom~A to Classroom~B halfway through the
year.  Alice's talent score is \emph{shared} between both classrooms.  The
per-kid regression knows this: it tracks Alice across both rooms and uses data
from \emph{both} rooms to estimate her talent.  But when you compute classroom
averages, each classroom computes its own average independently---Classroom~A
doesn't know that Alice also contributed to Classroom~B's average.  The
per-kid regression and the classroom-average regression ``see'' different
patterns, so they give different answers.

\subsection{Mathematical Explanation}

In the firm-level regression, $\gamma_f$ is a single parameter estimated using
\emph{all} observations of firm~$f$, regardless of which cells they belong to.
The Frisch--Waugh--Lovell (FWL) projector $M_\mathrm{firm}$ absorbs
$\{\gamma_f, \alpha_{mt}\}$ by projecting onto the orthogonal complement of the
column space of the fixed-effect design matrix $[D_f \,|\, D_{mt}]$.

\paragraph{When C3 is violated.}
Suppose firm $f$ operates in two cells, $c_1 = (j_1, m_1)$ and
$c_2 = (j_2, m_2)$.  Then $\gamma_f$ enters the normal equations for
\emph{both} cells, creating a \textbf{cross-cell link}.  When estimating
$\hat\gamma_f$, OLS uses firm~$f$'s observations from both $c_1$ and $c_2$.
This means the projected residual $\tilde{X}_{fmt}$ for an observation in $c_1$
depends on data from $c_2$.

After cell-averaging, the cell FEs are:
\[
  \bar\gamma_{c_1}
  = \frac{1}{N_{c_1}}\sum_{f' \in \cFpre_{c_1}} \gamma_{f'},
  \qquad
  \bar\gamma_{c_2}
  = \frac{1}{N_{c_2}}\sum_{f' \in \cFpre_{c_2}} \gamma_{f'}.
\]
Both contain $\gamma_f$ (since $f \in \cFpre_{c_1} \cap \cFpre_{c_2}$), so
they are \emph{not independent}.  However, the cell-level FWL projector
$M_\mathrm{cell}$ treats $\bar\gamma_{c_1}$ and $\bar\gamma_{c_2}$ as
\textbf{separate, independent parameters}: each cell's FE indicator column
$D_c$ is disjoint from every other cell's.  It does not ``know'' that
$\bar\gamma_{c_1}$ and $\bar\gamma_{c_2}$ share a component
$\gamma_f / N_{c_1}$ and $\gamma_f / N_{c_2}$, respectively.

As a result:
\[
  M_\mathrm{firm} \neq \text{``cell-averaged version of } M_\mathrm{cell}\text{''},
\]
which means the projected cell means differ between the two approaches:
\[
  \frac{1}{N_c}\sum_{f \in c} \widetilde{X}^{\,\mathrm{firm}}_{fmt}
  \;\neq\;
  \widetilde{\bar{X}}^{\,\mathrm{cell}}_{ct},
\]
and the normal equations no longer coincide even with $N_c$-weighting.

\paragraph{When C3 holds.}
If every firm belongs to exactly one cell, then $D_f$ is \emph{nested within}
$D_c$: all of firm~$f$'s observations lie in a single cell.  The firm FE
becomes a within-cell parameter.  Cross-cell links vanish, and the between-cell
components of the two projections agree.

\begin{example}
In the data, about 20\% of firms (accounting for 36\% of observations) operate
in more than one $(j, m)$ cell.  These multi-cell firms create exactly the
cross-cell links described above.
\end{example}

\bigskip
\hrule
\bigskip

%% ========================================================================
\section{Condition C5: Fixed Cell Composition Within Regime}
%% ========================================================================

\begin{condition}[C5 --- Fixed Cell Composition Within Regime]
Within each support regime (election cycle) $e$, the set of firms in each cell
is constant across all years $t$ of that regime:
\[
  \forall\, c = (j, m, e),\;\forall\, t, t' \in \text{years}(e):\quad
  \cFpre_c(t) = \cFpre_c(t').
\]
Equivalently, within each regime, every firm is present in \textbf{all} years of
the regime (a balanced panel within each regime).
\end{condition}

\subsection{Intuitive Explanation}

Imagine each classroom has a team roster that is supposed to be fixed for the
whole school year (one election cycle).  The classroom gets \emph{one name
tag}---the cell FE $\bar\gamma_c$---for the entire year.

But what if some kids join in March and others drop out in October?  The
classroom's ``average talent'' in September is different from its ``average
talent'' in February, because the team changed.  Yet we only have \emph{one}
name tag trying to represent the whole year.  It captures a sort of
``average of the averages,'' but it cannot track the month-to-month shifts.

The per-kid regression has no such problem: each kid has their own personal tag
that stays constant regardless of what other kids do.  Kids entering or leaving
the classroom doesn't confuse anything.

So if the team roster shifts within a year, the one-tag-per-classroom approach
cannot keep up, and the answers diverge.

\subsection{Mathematical Explanation}

From the averaging identity, the cell-level ``firm FE'' at time $t$ is:
\[
  \bar\gamma_c(t)
  = \frac{1}{|\cFpre_c(t)|}\sum_{f \in \cFpre_c(t)} \gamma_f,
\]
where $\cFpre_c(t)$ is the set of firms present in cell $c$ at time $t$.

\paragraph{When C5 holds.}
$\cFpre_c(t) = \cFpre_c$ is constant within regime $e$, so
$\bar\gamma_c(t) = \bar\gamma_c$ is a time-invariant constant (within the
regime).  A single FE per $(j, m, e)$ triple absorbs it perfectly.

\paragraph{When C5 is violated.}
If firms enter or exit within the regime, $\bar\gamma_c(t)$ varies by year.
The aggregated regression absorbs a single parameter $\delta_c$ per $(j, m, e)$
cell---effectively the OLS ``best constant approximation'' to
$\bar\gamma_c(t)$.  The residual
\[
  \eta_{ct} \equiv \bar\gamma_c(t) - \delta_c
\]
is time-varying and ends up in the error term $\epsilon_{ct}$.  If $\eta_{ct}$
correlates with $\overline{\FA}^{\ell,\mathrm{pre}}_{jmt}$ (which it may,
since both depend on \emph{which firms are present} at time $t$), the cell-level
OLS estimate $\hat\lambda^\mathrm{cell}$ is biased \emph{relative to}
$\hat\lambda^\mathrm{firm}$.

\begin{example}
Suppose cell $c$ has firms $\{A, B\}$ in year~1 and $\{A, B, C\}$ in year~2
(firm~$C$ enters in year~2).  Then:
\begin{align*}
  \bar\gamma_c(1) &= \tfrac{1}{2}(\gamma_A + \gamma_B), \\
  \bar\gamma_c(2) &= \tfrac{1}{3}(\gamma_A + \gamma_B + \gamma_C).
\end{align*}
The cell FE $\delta_c$ can absorb at most one of these, not both.  The leftover
$\eta_{c1}$ or $\eta_{c2}$ contaminates the error term.

Meanwhile, $\overline{\FA}_{c}(t)$ also shifts: in year~2, firm~$C$'s party
exposure enters the cell mean.  If firm~$C$ happens to be politically connected
(high $\FA$) \emph{and} has high $\gamma_C$ (unobservably productive), then
$\eta_{c2} > 0$ and $\overline{\FA}_{c}(2)$ is inflated, creating a spurious
positive correlation.

In the firm-level regression, each $\gamma_f$ is absorbed individually, so
firm~$C$'s entry creates no bias.
\end{example}

\bigskip
\hrule
\bigskip

%% ========================================================================
\section{Condition C6: No Within-Cell Regressor Heterogeneity}
%% ========================================================================

\begin{condition}[C6 --- No Within-Cell Regressor Heterogeneity After FE
Projection]
After absorbing fixed effects, the projected regressors are constant within
each cell-year:
\[
  \tilde{X}_{fmt} = \tilde{\bar{X}}_{ct}
  \quad \forall\, f \in \cFpre_c,\;\forall\, t,
\]
where tildes denote FE-projected residuals and
$\tilde{\bar{X}}_{ct} = N_c^{-1}\sum_{f \in c} \tilde{X}_{fmt}$.

Equivalently: \emph{within-cell, between-firm variation in the projected
$\FA^\ell$ regressors is zero.}
\end{condition}

\subsection{Intuitive Explanation}

Imagine two kids in the same classroom: Alice studies 5~hours and scores~90;
Bob studies 1~hour and scores~60.  The classroom average is ``3~hours studied,
75~points scored.''

The per-kid regression sees the \emph{within-classroom} pattern: the kid who
studied more scored more.  It uses this Alice-versus-Bob contrast to learn that
studying helps.

The classroom-average regression only sees \emph{one data point} for this
classroom: (3~hours, 75~points).  It cannot look \emph{inside} the classroom.
The within-classroom variation---the Alice-vs-Bob difference---is invisible to
it.

If both kids studied the same amount (say, both 3~hours, but Alice scored~90
and Bob scored~60 for other reasons), then the classroom average captures
\emph{all} the study-hour information perfectly---there is nothing hidden
inside.  Only then do the two approaches give the same answer.

This is why the synthetic test (which forces all ``kids'' in the same classroom
to have the same ``study hours'') works perfectly at $10^{-15}$ precision, but
real data (where firms in the same sector-municipality cell have different party
exposures) never will.

\subsection{Mathematical Explanation}

After absorbing fixed effects via FWL, the firm-level OLS normal equation is:
\begin{equation}\label{eq:firm_normal}
  \sum_c \sum_{f \in c}
  \tilde{X}_{fmt}'\bigl(\tilde{Y}_{fmt} - \tilde{X}_{fmt}\,\lambda\bigr) = 0.
\end{equation}

Decompose each observation into its cell-year mean and within-cell deviation:
\[
  \tilde{X}_{fmt}
  = \underbrace{\tilde{\bar{X}}_{ct}}_{\text{between-cell}}
  + \underbrace{(\tilde{X}_{fmt} - \tilde{\bar{X}}_{ct})}_{\text{within-cell}},
  \qquad
  \tilde{Y}_{fmt}
  = \tilde{\bar{Y}}_{ct}
  + (\tilde{Y}_{fmt} - \tilde{\bar{Y}}_{ct}).
\]

Substituting into~\eqref{eq:firm_normal} and expanding (using the fact that
within-cell deviations sum to zero), the normal equation splits into two
orthogonal components:
\begin{equation}\label{eq:decomp}
  \underbrace{
    \sum_c N_c\, \tilde{\bar{X}}_{ct}'
    \bigl(\tilde{\bar{Y}}_{ct} - \tilde{\bar{X}}_{ct}\,\lambda\bigr)
  }_{\displaystyle B(\lambda)\;\text{--- between-cell}}
  \;+\;
  \underbrace{
    \sum_c \sum_{f \in c}
    (\tilde{X}_{fmt} - \tilde{\bar{X}}_{ct})'
    \bigl(
      (\tilde{Y}_{fmt} - \tilde{\bar{Y}}_{ct})
      - (\tilde{X}_{fmt} - \tilde{\bar{X}}_{ct})\,\lambda
    \bigr)
  }_{\displaystyle W(\lambda)\;\text{--- within-cell}}
  = 0.
\end{equation}

The $N_c$-weighted cell-level OLS normal equation is \textbf{only the
between-cell part}:
\begin{equation}\label{eq:cell_normal}
  B(\lambda) = \sum_c N_c\, \tilde{\bar{X}}_{ct}'
  \bigl(\tilde{\bar{Y}}_{ct} - \tilde{\bar{X}}_{ct}\,\lambda\bigr) = 0.
\end{equation}

For both estimators to yield the same $\hat\lambda$, we need
\[
  B(\hat\lambda) + W(\hat\lambda) = 0
  \quad\text{and}\quad
  B(\hat\lambda) = 0
  \quad\Longrightarrow\quad
  W(\hat\lambda) = 0.
\]

\paragraph{When does $W(\hat\lambda) = 0$?}

Define the within-cell cross-product matrices:
\begin{align}
  W_{XX} &= \sum_c \sum_{f \in c}
    (\tilde{X}_{fmt} - \tilde{\bar{X}}_{ct})
    (\tilde{X}_{fmt} - \tilde{\bar{X}}_{ct})', \label{eq:Wxx} \\
  W_{XY} &= \sum_c \sum_{f \in c}
    (\tilde{X}_{fmt} - \tilde{\bar{X}}_{ct})
    (\tilde{Y}_{fmt} - \tilde{\bar{Y}}_{ct}). \label{eq:Wxy}
\end{align}

The within-cell normal equation $W(\lambda) = 0$ can be written as
$W_{XY} - W_{XX}\,\lambda = 0$.  This vanishes for all $\lambda$ if and only if
$W_{XX} = 0$ and $W_{XY} = 0$.  And $W_{XX} = 0$ requires:
\[
  \tilde{X}_{fmt} = \tilde{\bar{X}}_{ct}
  \quad\forall\, f \in c, \;\forall\, t.
\]
That is: after FE projection, every firm in the same cell-year must have the
\textbf{same} projected $\FA^\ell$ value.  In other words, there must be
\emph{zero within-cell variation} in the projected regressors.

\paragraph{Why this fails on real data.}

The $\FA^\ell_{fmt}$ instrument is the product of a firm-specific baseline party
exposure $\omega_{fp}$ and a municipality-level alignment indicator
$\mathrm{Align}^\ell_{mpt}$:
\[
  \FA^\ell_{fmt} = \sum_p \omega_{fp,0}\, \mathrm{Align}^\ell_{mpt}.
\]
The alignment term $\mathrm{Align}^\ell_{mpt}$ is common to all firms in the
same $(m, t)$, but the baseline exposure $\omega_{fp,0}$ varies across firms.
Firms in the same cell $(j, m, t)$ have different owners affiliated with
different parties, so $\omega_{fp,0}$ differs across $f$ within the cell.

After the firm FE projection (which removes each firm's time-mean), the
projected regressor is:
\[
  \tilde{\FA}^\ell_{fmt}
  = \FA^\ell_{fmt} - \bar{\FA}^\ell_{f\cdot}
  - \bar{\FA}^\ell_{\cdot mt} + \bar{\FA}^\ell_{\cdot\cdot},
\]
where $\bar{\FA}^\ell_{f\cdot}$ is firm $f$'s time-mean.  Since
$\omega_{fp,0}$ varies across firms, $\bar{\FA}^\ell_{f\cdot}$ also varies
across firms in the same cell, so $\tilde{\FA}^\ell_{fmt}$ differs across
firms within the cell.  Therefore $W_{XX} \neq 0$.

\paragraph{Magnitude of the gap.}

The firm-level $\hat\lambda^\mathrm{firm}$ solves
$B(\lambda) + W(\lambda) = 0$, while the cell-level
$\hat\lambda^\mathrm{cell}$ solves $B(\lambda) = 0$.  Using standard
between/within decomposition algebra:
\begin{align}
  \hat\lambda^\mathrm{firm}
  &= (B_{XX} + W_{XX})^{-1}(B_{XY} + W_{XY}), \label{eq:firm_est} \\
  \hat\lambda^\mathrm{cell}
  &= B_{XX}^{-1}\, B_{XY}, \label{eq:cell_est}
\end{align}
where $B_{XX} = \sum_c N_c\,\tilde{\bar{X}}_{ct}\tilde{\bar{X}}_{ct}'$ and
$B_{XY} = \sum_c N_c\,\tilde{\bar{X}}_{ct}\tilde{\bar{Y}}_{ct}$.

The gap is:
\[
  \hat\lambda^\mathrm{firm} - \hat\lambda^\mathrm{cell}
  = (B_{XX} + W_{XX})^{-1}(B_{XY} + W_{XY}) - B_{XX}^{-1} B_{XY}.
\]
This is zero if and only if $W_{XX} = 0$ and $W_{XY} = 0$.  On real data,
$W_{XX} / (B_{XX} + W_{XX})$ measures the \emph{share of total variation that
is within-cell}.  The larger this share, the larger the potential gap.

\begin{remark}[Why synthetic data achieves machine precision]
In the synthetic benchmark, $\FA^\ell_{fmt}$ is constructed to be constant
within each cell-year: $\FA^\ell_{fmt} = \FA^\ell_{ct}$ for all $f \in c$.
This forces $W_{XX} = 0$ and $W_{XY} = 0$, so
$\hat\lambda^\mathrm{firm} = \hat\lambda^\mathrm{cell}$ at machine precision
($\sim 10^{-15}$).
\end{remark}

\begin{remark}[C6 is the irreducible gap]
Conditions C1--C5 can all be enforced on real data (by reweighting, restricting
to single-cell firms, balancing the panel, using the correct FE, and disabling
singleton removal).  Condition~C6 \textbf{cannot} be enforced on real data
without modifying the instruments themselves, because within-cell heterogeneity
in $\omega_{fp,0}$ is an intrinsic feature of the data.  The residual gap after
enforcing C1--C5 (the ``Silver tier'' gap) is entirely attributable to~C6.
\end{remark}

\bigskip
\hrule
\bigskip

%% ========================================================================
\section{Summary}
%% ========================================================================

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{c p{3.5cm} p{4.5cm} p{3.5cm}}
\hline
\textbf{\#} & \textbf{Condition} & \textbf{What breaks without it} & \textbf{Enforceable on real data?} \\
\hline
C3 & Firm immobility (one cell per firm) &
  Firm FE creates cross-cell links that the cell-level projector ignores &
  Yes --- restrict to single-cell firms (drops $\sim$20\% of firms) \\[4pt]
C5 & Fixed composition within regime &
  Cell FE $\bar\gamma_c(t)$ becomes time-varying; single cell FE $\delta_c$
  leaves a correlated residual &
  Yes --- balance panel within regime (drops $\sim$40\% of single-cell obs) \\[4pt]
C6 & No within-cell regressor heterogeneity &
  Firm-level OLS uses within-cell variation ($W_{XX}, W_{XY}$) that cell-level
  OLS cannot see &
  \textbf{No} --- inherent to real data \\
\hline
\end{tabular}
\end{center}

\end{document}
