\documentclass[aspectratio=169,11pt]{beamer}
\usetheme{default}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{headline}{}
\setbeamertemplate{footline}{\hfill{\usebeamerfont{footline}\insertframenumber}\hspace{0.6em}\vspace{0.4em}}
\setbeamerfont{frametitle}{size=\large,series=\bfseries}
\setbeamerfont{footline}{size=\scriptsize}
\usepackage{palatino}
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{array}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{makecell}
\title{\textbf{Updated AR Test Results}}
\date{}
\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Headline date}
\centering
\vfill
{\Large May 14, 2026}
\vfill
\end{frame}

\begin{frame}{Overview}
\begin{itemize}\setlength{\itemsep}{0.45em}
\item \textbf{Question.} Under the updated instrument convention (Variant A muni-relative owner-share, pre-earliest-election window), does the Anderson-Rubin test reject $H_0\!:\!\beta=0$ on sector employment shares?
\item \textbf{Two taxonomies.} Policy block (Agriculture, Industry, Infrastructure, Services; $K=4$); firm-size (MPME, Media, Grande; $K=3$).
\item \textbf{Four control specs.} (1) No controls; (2) $+$ EC (per-cell exposure control); (3) $+$ Vol (volume control); (4) $+$ Vol $+$ EC.
\item \textbf{Four channels.} Mayor; Mayor $\cdot$ President; Mayor $\cdot$ Governor; Mayor $\cdot$ Gov.\ $\cdot$ President.
\item All regressions use municipality and year FE; one-way cluster on \texttt{muni\_id}.
\end{itemize}
\end{frame}

\input{slides_body_policy_block.tex}
\input{slides_body_size_bin.tex}

\begin{frame}{Bottom Line}
\begin{itemize}\setlength{\itemsep}{0.45em}
\item Under Variant A weights with pre-earliest-election windows, the AR test \textit{rejects} at 5\% in two cells: (M $\cdot$ G, $+$ EC) and (M $\cdot$ G, $+$ Vol $+$ EC) for policy\_block; (M $\cdot$ P, no controls / $+$ Vol) and (M $\cdot$ G, $+$ EC / $+$ Vol $+$ EC) for size\_bin.
\item The volume control (single coefficient) shifts $F$ negligibly; the EC controls (per-cell, $K-1$ columns) move $F$ in both directions.
\item Size\_bin power is constrained ($K-1=2$ instruments per channel).
\item Per-channel coefficients reported on the following slides for inspection.
\end{itemize}
\end{frame}

\end{document}

