\begin{frame}{Firm Size (S3) --- Setup}
\begin{itemize}\setlength{\itemsep}{0.4em}
\item \textbf{Taxonomy:} Firm Size (S3). $K=3$; hold-out sector: \textbf{Grande}.
\item \textbf{Levels:} MPME, Media, Grande.
\item Weights: Variant A (muni-relative aligned-owner share), pre-earliest-election window.
\item Volume control: $\mathrm{vol}_{mt} = \mathrm{total\_bndes\_real}_{mt} / \mathrm{pib\_real}_{m,2002}$.
\item FE: municipality + year. SE: one-way cluster on \texttt{muni\_id}.
\end{itemize}
\end{frame}

\begin{frame}{Firm Size (S3) --- AR joint $F$}
{\footnotesize Variant A weights, pre-earliest-election window.}
\vskip 0.25em
\resizebox{\textwidth}{!}{%
\setlength{\tabcolsep}{4pt}%
\footnotesize%
\input{tables/ar_table_fstats_size_bin.tex}%
}
\vskip 0.5em
{\footnotesize Cluster-robust AR Wald joint $F$. 
$^{*}$ $p<0.10$, $^{**}$ $p<0.05$, $^{***}$ $p<0.01$. 
[$p$] is the $p$-value; reject at 5\% iff $p<0.05$.}
\end{frame}

\begin{frame}{Firm Size (S3) --- Coefficients: Mayor}
\input{tables/ar_table_coefs_size_bin_M.tex}
\vskip 0.25em
{\footnotesize SE in parentheses. EC (mean) row reports the mean coefficient across the $K-1$ EC controls. Stars: $^{*}$ $p<0.10$, $^{**}$ $p<0.05$, $^{***}$ $p<0.01$.}
\end{frame}

\begin{frame}{Firm Size (S3) --- Coefficients: Mayor $\cdot$ President}
\input{tables/ar_table_coefs_size_bin_MP.tex}
\vskip 0.25em
{\footnotesize SE in parentheses. EC (mean) row reports the mean coefficient across the $K-1$ EC controls.}
\end{frame}

\begin{frame}{Firm Size (S3) --- Coefficients: Mayor $\cdot$ Governor}
\input{tables/ar_table_coefs_size_bin_MG.tex}
\vskip 0.25em
{\footnotesize SE in parentheses. EC (mean) row reports the mean coefficient across the $K-1$ EC controls.}
\end{frame}

\begin{frame}{Firm Size (S3) --- Coefficients: Mayor $\cdot$ Gov. $\cdot$ President}
\input{tables/ar_table_coefs_size_bin_MGP.tex}
\vskip 0.25em
{\footnotesize SE in parentheses. EC (mean) row reports the mean coefficient across the $K-1$ EC controls.}
\end{frame}

